---
title: "feb-02-analysis"
format: html
editor: visual
---

## Bills Introduced in the 2025 Texas Legislative Session

### Up to February 2

This is the analysis portion of bill data since the start of the filing period up until Feb. 2. I'm importing the data file I saved from the cleaning portion. These are some of the questions I hope to answer:

-   Which lawmakers are filing the most? And how does that relate to party?

-   Where are the bills coming from across the State?

-   How many bills are associated with different policy areas? Examples being abortion, immigration, environment, education, religion, insurance, etc. (will use grepyl to code for key words in strings)

-   Look at who is filing the most bills per seat ... so divide number of bills per party by number of seats for that party

### Setup

```{r}
library(tidyverse)
library(janitor)
```

### Importing my clean data

```{r}
bill_data <- read_rds("data-processed/02texasbills.rds")

bill_data
```

### Filtering bill description by keyword

For example, how many bills use the term abortion?

```{r}
bill_data |> 
  filter(grepl("abortion", description, ignore.case = TRUE))
```

What about terms related to abortion?

```{r}
#trying to filter description by keyword
bill_data |> 
  filter(grepl("fetal|viability|pregnancy|termination|miscarriage|minor|consent|rape|incest|remains|patient|privacy", description, ignore.case = TRUE))
```

How many bills use the words diversity, equity, inclusion, or DEI?

```{r}
bill_data |> 
  filter(grepl("diversity|equity|inclusion|DEI", description, ignore.case = TRUE))
```

How many bills mention the word immigration in the description?

```{r}
bill_data |> 
  filter(grepl("immigration", description, ignore.case = TRUE))
```

What about terms relating to immigration?

```{r}
bill_data |> 
  filter(grepl("border|asylum|refugee|citizenship|deportation|naturalization|undocumented|detention|dreamers", description, ignore.case = TRUE))
```

How many bills use the words environment or energy in the description?

```{r}
bill_data |> 
  filter(grepl("environment|energy", description, ignore.case = TRUE))
```

What about terms relating to environment/energy?

```{r}
bill_data |> 
  filter(grepl("conservation|water|natural|gas|power|grid|climate|greenhouse|emissions|air|pollution|renewable|fossil|fuel|offshore|fracking|sustainability", description, ignore.case = TRUE))

```

I don't understand all of these results.

How many bills use the term vouchers in the description?

```{r}
bill_data |> 
  filter(grepl("voucher|vouchers", description, ignore.case = TRUE))
```

How many bills use terms related to school choice?

```{r}
bill_data |> 
  filter(grepl("savings|account|school|choice|parental|education|K-12", description, ignore.case = TRUE))
```

How many bills use the term prayer in the description?

```{r}

```

### Finding lawmakers who filed most vs. least bills

```{r}
#lawmakers who filed most bills 
bill_data |> 
  select(name, party, role, district) |> 
  count(name, party, district) |> 
  arrange(desc(n))
```

```{r}
#lawmakers who filed least bills 
bill_data |> 
  select(name, party, role, district) |> 
  count(name, party, district) |> 
  arrange(n)
```

```{r}
#spliting this up by party, democrats first
bill_data |> 
  select(name, party, role, district) |> 
  group_by(party) |> 
  count(name) |> 
  filter(party == "D", n > 10) |> 
  arrange(desc(n))
```

```{r}
#now republicans
bill_data |> 
  select(name, party, role, district) |> 
  group_by(party) |> 
  count(name) |> 
  filter(party == "R", n > 10) |> 
  arrange(desc(n))
```

### Which party has filed the most bills?

```{r}
bill_data |> 
  select(party) |> 
  count(party)
```

### What types of lawmakers are filing the most bills?

```{r}
bill_data |> 
  group_by(party) |> 
  count(role) |> 
  arrange(desc(n))
```

Is this proportional to seats per party?

-   Republican seats in House is 88

-   Democrat seats in House is 62

-   Republican seats in Senate is 20

-   Democrat seats in Senate is 11

```{r}

```
