---
title: "01-cleaning"
format: html
editor: visual
---

## Bills introduced in the 2025 Texas Legislative Session

### Up to February 2

This is a continuation of my analysis of bills filed in the current Texas Legislative session. This document shows my cleaning of the data. I'm combining and cleaning data from three different CSVs to include bill_id, bill_number, date_filed, and description as well as important identifying information for the sponsor of each bill.

### Setup

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```

### Importing the data

```{r}
#individual bills
bills <- read_csv("data-raw/bills.csv") 

bills
```

```{r}
#individual people/lawmakers
lawmakers <- read_csv("data-raw/people.csv")

lawmakers
```

```{r}
#sponsors associated with each bill 
sponsors <- read_csv("data-raw/sponsors.csv")

sponsors
```

### Cleaning up each table

```{r}
#bill details 
bills_clean <- bills |> 
  rename(date_filed = status_date) |> 
  select(-session_id, -status, -status_desc, -title, -committee_id, -committee, -last_action_date, -last_action, -url, -state_link) |> 
  filter(date_filed > "2025-01-17")

bills_clean
```

```{r}
#lawmaker details
lawmakers_clean <- lawmakers |> 
  select(people_id, name, party, role, district) 

lawmakers_clean
```

Note: I don't really need to do anything to the sponsor table

### Merging the data together to make one table

```{r}
#lawmaker and sponsor data first
lawmaker_sponsor <- merge(lawmakers_clean, sponsors) |> 
  select(-position)

lawmaker_sponsor
```

```{r}
#adding date filed and description 
bill_details <- lawmaker_sponsor |> 
  left_join(bills_clean, by = join_by(bill_id)) |> 
  filter(date_filed > "2025-01-17")

bill_details
```

I'm not sure why there are gaps in some of these.

### Saving this table for my analysis

```{r}
#writing an rds file
bill_details |> 
  write_rds("data-processed/02texasbills.rds")
```
