---
title: "TX Lege Analysis, practice"
output: html_notebook
---

This is an analysis of bills introduced between the start of filing up until
January 17. I'm still trying to figure out exactly how I want to do this, so
this is mainly me experimenting with the dataset. I'm using data from LegiScan
to look at all bills introduced to the session so far. 

## Setup 
```{r}
library(tidyverse)
library(janitor)
```
## Importing bill data 

```{r}
new_bills <- read_csv("Lege-Jan17/data-raw/bills.csv")

new_bills
```

```{r}
bill_details <- new_bills |> 
  rename(date_filed = status_date) |> 
  select(-session_id, -status, -status_desc, -title, -committee_id, -committee, -last_action_date, -last_action, -url, -state_link)

bill_details
```

## Importing sponsor data for each bill 

```{r}
sponsor_IDs <- read_csv("Lege-Jan17/data-raw/sponsors.csv")

sponsor_IDs
```

## Importing lawmaker data

```{r}
lawmakers <- read_csv("Lege-Jan17/data-raw/people.csv")

lawmakers
```

```{r}
#selecting the columns I want
lawmakers_clean <- lawmakers |> 
  select(people_id, name, party, role, district)

lawmakers_clean
```

## Merging the sponsor and lawmaker datasets

```{r}
bill_IDs <- merge(lawmakers_clean, bill_sponsors) |> 
  select(-position)

bill_IDs
```

## Merging all of the data together 

#I want to add the date_filed and description columns from my bill_details data table 

```{r
```



