---
title: "01-cleaning-NEW"
format: html
editor: visual
---

## Bills Introduced in the 2025 Texas Legislative Session

### November 12 to January 17

This is the start of my analysis of bills filed in the current Texas Legislative session. This document shows my cleaning of the data. I'm combining and cleaning data from three different csvs to include bill_id, bill_number, date_filed, and description as well as important identifying information for the sponsor of each bill.

### Setup

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```

### Importing the Data

```{r}
#individual bills 
bills <- read_csv("data-raw/bills.csv")

bills
```

```{r}
#individual people/lawmakers
lawmakers <- read_csv("data-raw/people.csv")

lawmakers
```

```{r}
#sponsors for each bill
sponsors <- read_csv("data-raw/sponsors.csv")

sponsors
```

### Cleaning up each table

```{r}
#bill details 
bills_clean <- bills |> 
  rename(date_filed = status_date) |> 
  select(-session_id, -status, -status_desc, -title, -committee_id, -committee, -last_action_date, -last_action, -url, -state_link) 

bills_clean
```

```{r}
#lawmaker details
lawmakers_clean <- lawmakers |> 
  select(people_id, name, party, role, district)

lawmakers_clean
```

Note: the sponsor data is fine as is

### Merging the data to make one big table

```{r}
#lawmaker and sponsor data first
lawmaker_sponsor <- merge(lawmakers_clean, sponsors) |> 
  select(-position)

lawmaker_sponsor
```

```{r}
#adding date filed and description columns
bill_details <- lawmaker_sponsor |> 
  left_join(bills_clean, by = join_by(bill_id)) 

bill_details
```

### Saving this table for my analysis

```{r}
#writing an rds file
bill_details |> 
  write_rds("data-processed/01-texasbills.rds")

```
